Augmentation of a Model Historical Harpsichord Mechanism for Haptic Enabled Interaction

### 2025-01-18 src/article.tex
introduction wording

Some items in the Tagliavini collection are no longer in playing condition or
are too fragile to continue to be available to museum visitors. One of those
fragile instruments is a 1547 Harpsichord in the Italian style by Alessandro
Trasuntino \cite{Wraight2024}\footnote{examples of similar instruments by
Trasuntino available online in the Royal College of Music Collection (1531:
\url{https://museumcollections.rcm.ac.uk/collection/Details/collect/58}) and the
Musée de la musique collection (1538:
\url{https://carmentis.kmkg-mrah.be:443/eMP/eMuseumPlus?service=ExternalInterface&module=collection&objectId=106088&viewType=detailView})
}, which was used as the basis for an interactive exhibition organised by San
Colombano.

@@

The "Haptic Key" project by Timmermans et al. \cite{Timmermans2020} began with
the model of a piano key. A 3-key cross-sectional model of a harpischord
mechanism \ref{fig:3key} was provided to the NEMUS project by San Colombano as a
test bed for sensors and as proof of concept for a larger scale interface. Each
key of the model has 2 sets of strings that are under tension but not at any
specific musical pitch. The strings are there to provide the same resistance to
the jack plectra as would be found on an instrument. The mechanism is in the
Italian style of the XVI-XIX century with 2 jacks per key and their
corresponding strings tuned in unison \footnote{"8'-8' È la dispozione normale
dei clavicembali italiani" \cite{Tagliavini1987}}.

@@

McAlpine outlines a similar situation to the Tagliavini collection in
his work with the Benton Fletcher Collection \cite{McAlpine2014}. There
was a stipulation by Benton Fletcher that the instruments remain
available to play when they were gifted to National Trust Fenton House.
A custom MIDI interface arranged in a two manual harpischord style which
triggered samples of the instruments in the collection using an
appropraite recording strategy for each. One problem highlighted in user
tests by McAlpine is that the weighted keys did no provide ``an
experiential sense of interacting with a historical keyboard''
\cite{McAlpine2014}. The MIDI interface created from commercially
available weighted keys. McAlpine posits a haptic keyboard augemented
with ``actuators to provide positionally-sensitive real-time force
feedback at point-of-contact'' in a similar approach as the piano
mechanism by Gillespie \cite{Gillespie1996}.

>>>>

Some items in the Tagliavini collection are no longer in playing condition or
are too fragile to continue to be available to museum visitors. One of those
fragile instruments is a 1547 Harpsichord in the Italian style by Alessandro
Trasuntino \cite{Wraight2024}\footnote{examples of similar instruments by
Trasuntino available online in the Royal College of Music Collection (1531:
\url{https://museumcollections.rcm.ac.uk/collection/Details/collect/58}) and the
Musée de la musique collection (1538:
\url{https://carmentis.kmkg-mrah.be/eMP/eMuseumPlus?service=ExternalInterface&objectId=106088})
}, which was used as the basis for an interactive exhibition organised by San
Colombano.

@@

The "Haptic Key" project by Timmermans et al. \cite{Timmermans2020} began with
the model of a piano key. A 3-key cross-sectional model of a harpsichord
mechanism \ref{fig:3key} was provided to the NEMUS project by San Colombano as a
test bed for sensors and as proof of concept for a larger scale interface. Each
key of the model has 2 sets of strings that are under tension but not at any
specific musical pitch. The strings are there to provide the same resistance to
the jack plectra as would be found on an instrument. The mechanism is in the
Italian style of the XVI-XIX century with 2 jacks per key and their
corresponding strings tuned in unison \footnote{``8'-8' È la disposizione normale
dei clavicembali italiani'' \cite{Tagliavini1987}}.

@@

McAlpine outlines a similar situation to the Tagliavini collection in
his work with the Benton Fletcher Collection \cite{McAlpine2014}. There
was a stipulation by Benton Fletcher that the instruments remain
available to play when they were gifted to National Trust Fenton House.
A custom MIDI interface arranged in a two manual harpsichord style which
triggered samples of the instruments in the collection using an
appropraite recording strategy for each. One problem highlighted in user
tests by McAlpine is that the weighted keys did no provide ``an
experiential sense of interacting with a historical keyboard''
\cite{McAlpine2014}. The MIDI interface created from commercially
available weighted keys. McAlpine posits a haptic keyboard augemented
with ``actuators to provide positionally-sensitive real-time force
feedback at point-of-contact'' in a similar approach as the piano
mechanism by Gillespie \cite{Gillespie1996}.

### 2025-01-18 src/main.tex
introduction wording

\author{Roberto Livi}
\author{Catalina Vincens}
\affiliation{%
  \institution{Museo San Colombano}
  \city{Bologna}
  \country{Italy}}

\author{Andrew McPherson}
\affiliation{%
 \institution{Imperial College}
 \city{London}
 \country{United Kingdom}}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
This work describes the design of an electronically augmented replica of a
16th-century harpsichord keyboard with a typical Renaissance Italian layout. The
keyboard is a collaboration between the authors and the Museo di San Colombano,
which houses the Tagliavini Collection of playable historical keyboard
instruments. The keyboard was commissioned for exhibition purposes and to
enhance the experiential authenticity of the collection. The keyboard emulates
the mechanics of a two-register harpsichord featuring a single manual. Unpitched
strings under tension replicate the mechanical resistance characteristic of the
original compass. An array of infrared LED/phototransistor pairs, connected to a
microcontroller, was installed within the interface to generate a continuous
stream of data corresponding to the jacks' motion. This paper details the design
principles, the development, open source implementation, and execution of
the installation for the San Colombano exhibition.
\end{abstract}

@@

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{images/49-key-front.png}
  \caption{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
  \Description{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
  \label{fig:teaser}
\end{teaserfigure}

% \begin{teaserfigure}
%   \includegraphics[width=\textwidth]{images/side_w-o_sensors.png}
%   \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023}
%   \Description{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023}
%   \label{fig:teaser}
% \end{teaserfigure}

@@

\include{article.tex}

>>>>

\author{Roberto Livi}
\author{Catalina Vincens}
\affiliation{%
  \institution{Museo di San Colombano}
  \city{Bologna}
  \country{Italy}}

\author{Andrew McPherson}
\affiliation{%
 \institution{Imperial College London}
 \city{London}
 \country{United Kingdom}}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
% \renewcommand{\shortauthors}{Hamilton et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
This work describes the design of an electronically augmented replica of a
16th-century harpsichord keyboard with a typical Renaissance Italian layout. The
keyboard is a collaboration between the authors and the Museo di San Colombano,
which houses the Tagliavini Collection of playable historical keyboard
instruments. The keyboard was commissioned for exhibition purposes and to
enhance the experiential authenticity of the collection. The keyboard emulates
the mechanics of a two-register harpsichord featuring a single manual. Unpitched
strings under tension replicate the mechanical resistance characteristic of the
original compass. An array of infrared LED / phototransistor pairs, connected to a
microcontroller, was installed within the interface to generate a continuous
stream of data corresponding to the jacks' motion. This paper details the design
principles, the development, open source implementation, and execution of
the installation for the San Colombano exhibition.
\end{abstract}

@@

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.

% % Option 1:
% \begin{teaserfigure}
%   \includegraphics[width=\textwidth]{images/49-key-front.png}
%   \caption{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
%   \Description{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
%   \label{fig:teaser}
% \end{teaserfigure}

% Option 2:
\begin{teaserfigure}
 \hspace*{0.6in}
  \includegraphics[width=0.8\textwidth]{src/images/49-key-front-top.png}
  \caption{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
  \Description{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
  \label{fig:teaser}
\end{teaserfigure}

% % Option 3:
% \begin{teaserfigure}
%   \includegraphics[width=\textwidth]{src/images/sensors-without-keys-front.jpg}
%   \caption{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
%   \Description{49-key Model Harpsichord Mechanism by Roberto Livi, Bologna 2024}
%   \label{fig:teaser}
% \end{teaserfigure}

%% Placeholder
% \begin{teaserfigure}
%   \includegraphics[width=\textwidth]{images/side_w-o_sensors.png}
%   \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023}
%   \Description{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023}
%   \label{fig:teaser}
% \end{teaserfigure}

@@

\include{src/article}
